# Test data

## Peat depth and peatland delineation

To create the methodology for mapping peatland depth profiles, and estimating peat volume, carbon stocks and sensitivity to sampling effort, I will use two contrasting test sites: _Tydal_ and _Geilo._ Later, in chapter \@ref(additional-test-sites) I will validate the generality of this method on an additional four sites. For more details on the justification and the methods used, please the the manuscript.

Libraries used:
```{r, message=F, warning=F}
library(tmap)
library(sf)
library(readr)
library(tmaptools)
library(basemaps)
library(ggplot2)
library(ggpubr)
library(gstat)
library(matrixStats)
library(ggtext)
library(tidyverse)
library(osmplotr)
```

Import shape files with peatland delineations and fix the CRS.

```{r, message=F, warning=F}
SHP_tydal <- sf::read_sf("Data/Tydal/stasjon_Setermyra.shp")
SHP_geilo  <- sf::read_sf("Data/Geilo/geilo-dybdef.shp")

#st_crs(SHP_tydal)
#st_crs(SHP_geilo)# NA
# I found CRS through trial and error. It is UTM 32
st_crs(SHP_geilo) <- 25832 

# Transform to UTM33N
SHP_geilo <- st_transform(SHP_geilo, 25833)
```

Import peat depth measurements for the two sites.

```{r import depth data, message=F, warning=F}
depths_tydal <- readr::read_csv("Data/Tydal/Torvdybder_Tydal_stasjon.csv")
depths_geilo <- read.csv("Data/Geilo/torvdybder.csv", sep=";")


# convert data.frames to simple features.  
depths_tydal <- sf::st_as_sf(x = depths_tydal, 
                        coords = c("coords.x1", "coords.x2"),
                        crs = "+init=epsg:25833")
depths_geilo <- st_as_sf(x = depths_geilo, 
                        coords = c("x", "y"),
                        crs = "+init=epsg:25832")

# Tranbsform to UTM33, same as the shape file
depths_geilo <- st_transform(depths_geilo, 25833)
```

```{r, fig.cap="Confirming overlap between shape file and depth measurements"}
# Confirm overlap. Using base plotting to avoid 
# auto-transformation
plot(SHP_geilo$geometry,
     col="red"
     )
plot(depths_geilo$geometry, 
     pch = 16,
     col="grey",
     cex=.1,
     add=T)

```

That looks fine.

Download basemaps for some context.

```{r, warning=F, message=F}
# osmplotr want bboxes in latlong
SHP_geilo_ll <- sf::st_transform(SHP_geilo, 4326)
SHP_tydal_ll <- sf::st_transform(SHP_tydal, 4326)

bb_Geilo <- sf::st_bbox(SHP_geilo_ll)
bb_Tydal <- sf::st_bbox(SHP_tydal_ll)

# GEILO
base_geilo_hw <- 
  osmplotr::extract_osm_objects(bbox = bb_Geilo,
                                            key = c("highway"),
                                            sf = T)
base_geilo_building <- 
  osmplotr::extract_osm_objects(bbox = bb_Geilo,
                                            key = c("building"),
                                            sf = T)
base_geilo_ww <- 
  osmplotr::extract_osm_objects(bbox = bb_Geilo,
                                            key = c("waterway"),
                                            sf = T,
                                            return_type = "line")

# TYDAL
base_tydal_hw <- 
  osmplotr::extract_osm_objects(bbox = bb_Tydal,
                                            key = "highway",
                                            sf = T)
base_tydal_building <- 
  osmplotr::extract_osm_objects(bbox = bb_Tydal,
                                            key = c("building"),
                                            sf = T)
base_tydal_ww <- 
  osmplotr::extract_osm_objects(bbox = bb_Tydal,
                                            key = c("waterway"),
                                            sf = T,
                                            return_type = "line")

```

Plot static map

```{r, fig.cap="Map of Geilo and Tydal test sites. The peatland is deliminated as a green polygon(s). Dots are depth measurements and for Tydal the colour and size of the dots reflects the measured peat depths in meters."}

# GEILO
static_geilo <- tm_shape(SHP_geilo_ll)+
  tm_polygons(col="green")+
tm_shape(depths_geilo)+
  tm_dots()+
tm_shape(base_geilo_hw)+
  tm_lines()+
tm_shape(base_geilo_ww)+
  tm_lines(col="blue",
           size=2)+
tm_shape(base_geilo_building)+
  tm_polygons(col="black",
              alpha = .3)+
tm_compass()+
tm_scale_bar()+
  tm_layout(title="Geilo")


# TYDAL
static_tydal <- tm_shape(SHP_tydal_ll)+
  tm_polygons(col="green")+
tm_shape(depths_tydal)+
  tm_dots(size="Dybde",
          col="Dybde",
          palette="-viridis")+
tm_shape(base_tydal_ww)+
  tm_lines(col="blue",
           size=2)+
tm_compass()+
tm_scale_bar()+
tm_layout(title="Tydal",
          legend.show = F,
            inner.margins = c(.1,.02,.1,.02))

tmap_arrange(static_geilo,
             static_tydal)
```

These to test cases are very different. Geilo is a set of several unique mire polygons. Usually, in development projects, one would estimate the peat volume and C stock for each of these separately, but we will try now to see if it can be done in one operation. Tydal is a more typical example of a clear peatland delineation. Both cases have dense peat depth measurements taken.
