# Peat volume
We want to intrapolate values from the depth measurement only a raster grid. From this we can estimate the total volume of the peatland, and also visualise the depth profiles.

## Raster grid
Create raster grid based on the extent of the peatland
```{r}
#Create empty raster grid (stars object) based on the extent of the peatland shape files. Resolution 1 = 1 m (becasue crs is UTM)
grid_Tydal_stars <- starsExtra::make_grid(SHP_tydal, 1)
grid_Geilo_stars <- starsExtra::make_grid(SHP_geilo, 1)
```

Visualise the grid
```{r, fig.cap="A 1x1 m raster grid in grey, overlayed with the peatland delineation in yellow."}
# To plot the grid we can convert the stars objects to sf. That way we can plot as polygons and visualise the individual cells.
# This takes too long to perform for Geilo, but we can do it for Tydal:
grid_Tydal_sf <- st_as_sf(grid_Tydal_stars)

tm_shape(grid_Tydal_sf)+
  tm_polygons()+
  tm_shape(SHP_tydal)+
  tm_polygons(col="yellow",
              alpha=.5)
```

```{r, fig.cap="Preview of the cropped raster grids.", warning=F, message=F}
grid_Tydal_stars_crop <- sf::st_crop(grid_Tydal_stars, SHP_tydal)
grid_Geilo_stars_crop <- sf::st_crop(grid_Geilo_stars, SHP_geilo)

tmap_arrange(
tm_shape(grid_Geilo_stars_crop)+
  tm_raster(palette = "blue")+
  tm_layout(title = "Geilo",
            legend.show = F),
tm_shape(grid_Tydal_stars_crop)+
  tm_raster(palette = "red")+
  tm_layout(title = "Tydal",
            legend.show = F))

```

## Inverse distance weighting
We chose to use IDW for the interpolation. This is because we do not want to include a separate step with for example calculating a variogram. 

The IDW requires setting the number of neighbouring points to consider. We set this to consider all points in the dataset.
IDW also has a power parameter, and this is where we can tune the results by down-weighting datapoints that are further away.
Let's do a test with power equals 2 and power equals 3.
```{r, message=F}
library(gstat)

myPowers <- c(0:10)
for(i in myPowers){
  temp <- gstat::idw(formula = Dybde ~ 1, 
           locations = depths_tydal, 
           newdata = grid_Tydal_stars_crop, 
           idp=i)
  assign(paste0("IDW_tydal_", i), temp)
}

for(i in myPowers){
  temp <- gstat::idw(formula = Dybde ~ 1, 
           locations = depths_tydal, 
           newdata = grid_Geilo_stars_crop, 
           idp=i)
  assign(paste0("IDW_geilo_", i), temp)
}
```

What do these models look like. We will only look at Tydal, and not Geilo.
```{r, fig.cap="First view of how IDW models look like depending on the power parameter. Test site: Tydal.", message=F}
for(i in myPowers){
  
  temp <- get(paste0("IDW_tydal_", i))
  
  temp <- tm_shape(temp)+
    tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Interpolated peat\ndepth (m)")+
    tm_layout(legend.outside = T,
            title = paste0("Power = ", i))
  assign(paste0("tmap_IDW_tydal_power_", i), temp)
}



tmap_arrange(
  tmap_IDW_tydal_power_0,
  tmap_IDW_tydal_power_1,
  tmap_IDW_tydal_power_2,
  tmap_IDW_tydal_power_4,
  tmap_IDW_tydal_power_6,
  tmap_IDW_tydal_power_8,
ncol = 3)

```

A low power creates a more flat map where all the depth measurements count equally to any predicted value. With power set to 0 we just predict the average of all the measurened peat depths. For this case, increasing the power beyond 4 has little effect.

Let's check the fit with a plot of the residuals.
```{r}
# Prepare data frame with the measured data
resids_tydal <- data.frame(ID = 1:nrow(depths_tydal))
resids_geilo <- data.frame(ID = 1:nrow(depths_geilo))

resids_tydal$measured <- depths_tydal$Dybde
resids_geilo$measured <- depths_geilo$Dybde

# Get the predicted values
for(i in myPowers){
  temp <- get(paste0("IDW_tydal_", i))
  fit  <- stars::st_extract(temp, depths_tydal)
  resids_tydal[,i+3] <- resids_tydal$measured-fit$var1.pred
}

for(i in myPowers){
  temp <- get(paste0("IDW_geilo_", i))
  fit  <- stars::st_extract(temp, depths_geilo)
  resids_geilo[,i+3] <- resids_geilo$measured-fit$var1.pred
}

# Lots of NA's produced for depth measurements taken outside the peatland polygons
#tmap_mode("plot")
#tm_shape(IDW_geilo_10)+
#  tm_raster(col="var1.pred")+
#  tm_shape(fit[is.na(fit$var1.pred),])+
#  tm_symbols(size = .1,
#             col="yellow")

resids_geilo <- resids_geilo[!is.na(resids_geilo$V3),]


resids_tydal <- tidyr::pivot_longer(resids_tydal,
                                    !ID&!measured,
                                    names_to = "power",
                                    values_to = "residual")
resids_geilo <- tidyr::pivot_longer(resids_geilo,
                                    !ID&!measured,
                                    names_to = "power",
                                    values_to = "residual")


# Remove V from column names
resids_tydal$power <- gsub(pattern = "V", 
                           replacement = "",
                           x = resids_tydal$power)
resids_geilo$power <- gsub(pattern = "V", 
                           replacement = "",
                           x = resids_geilo$power)

resids_tydal$power <- as.numeric(resids_tydal$power)
resids_geilo$power <- as.numeric(resids_geilo$power)

resids_tydal$power <- resids_tydal$power-3
resids_geilo$power <- resids_geilo$power-3


gg_resids_tydal <- 
  ggplot(resids_tydal, aes(x=factor(power), y = residual))+
  geom_boxplot(fill="grey",
              trim=F)+
  theme_bw(base_size = 12)+
  xlab("Power")+
  ggtitle("Tydal")


gg_resids_geilo <- 
  ggplot(resids_geilo, aes(x=factor(power), y = residual))+
  geom_boxplot(fill="grey",
              trim=F)+
  theme_bw(base_size = 12)+
  xlab("Power")+
  ggtitle("Geilo")

ggpubr::ggarrange(gg_resids_geilo,
                  gg_resids_tydal)
```


```{r}


mean(depths_geilo$Dybde)
temp <- IDW_geilo_6
fit  <- stars::st_extract(temp, depths_geilo)
mean(fit$var1.pred, na.rm = T)
```

With a power of 2 or greater, the residuals are very stable. 

What about the estimated volume - does that change?
```{r}
volume_tydal <- NULL
for(i in myPowers){
  temp <- get(paste0("IDW_tydal_", i))
  volume_tydal <- c(volume_tydal,
                    sum(temp$var1.pred, na.rm = T))
}

```

```{r, fig.cap="Peat volume estimates as a function of the power parameter."}
volume_tydal <- data.frame("volume" = volume_tydal,
                              "power" = myPowers)
volume_tydal$relative_volume <- volume_tydal$volume/mean(volume_tydal$volume)*100
ggplot(volume_tydal, aes(x = factor(power), y = relative_volume))+
  geom_point(size=10)+
  xlab("power")+
  ylab("Peat volume as a percentage of\nmean predicted peat volume")+
  theme_bw(base_size = 12)
```

The values are stabilizing. In general though, the difference between the worst and best models is just 2-3%. 
This relationship probably depends a bit on the number of data points and the density of peat depth measurements. So we should not generalise just yet.

Now let's do a proper test of model fit, using leave one out cross validation.
```{r}
MAE_tydal <- data.frame(power = myPowers,
                        MAE = as.numeric(NA))
for(i in myPowers){
  temp <- krige.cv(Dybde ~ 1, depths_tydal, set = list(idp=i))
  MAE_tydal$MAE[MAE_tydal$power==i] <- mean(abs(temp$residual))
}

MAE_tydal$best <- ifelse(MAE_tydal$MAE==min(MAE_tydal$MAE), "best", "not-best")
```

```{r, fig.cap="Mean absolute error based on leave one out cross validation. The optimal power setting is indikcated by a green star symbol."}
ggplot(MAE_tydal, aes(x = power, y = MAE,
                      colour = best,
                      shape = best))+
  geom_point(size=10)+
  theme_bw(base_size = 12)+
  scale_x_continuous(breaks = 0:10)+
  guides(colour="none",
         shape = "none")+
  scale_color_manual(values = c("darkgreen","grey"))+
  scale_shape_manual(values = c(18, 19))

```


