# Validate using reduced sampling data

In chapter 2 we found the optimal power settings for the Tydal and Geilo sites. Now I just want to look at how this might have changed when we have removed *superflous* data points as in chapter 3.

Lets say for Tydal that we want to have 20 meters between the sampling points.

```{r, fig.cap="Peat depth measurements in the Tydal test site, after trimming the data poins so that the median distance between them is 20 meters."}
tm_shape(SHP_tydal)+
  tm_polygons()+
tm_shape(depths_tydal_i60)+
  tm_dots(size="Dybde",
          scale=2,
          col = "Dybde",
          palette = "-viridis",
          title = "Measured peat depth (m)")+
  tm_layout(legend.outside = T)
```
Confirm the median distance is 20 meters
```{r}
temp <- sf::st_distance(depths_tydal_i60) 
temp <- units::drop_units(temp)
temp[temp == 0] <- NA
temp <- rowMins(temp, na.rm=T)
median(temp)
```

Then we determine the optimal power setting.

```{r}
ccalc_optimumPower <- function(powerRange = 1:6,
                               nmax = 20,
                               peatDepths,
                               peatlandDelimination,
                               title
                               ){
  
  temp <- data.frame(power = powerRange,
                     MAE = as.numeric(NA))
  
  vol <- NULL
  
  myGrid <- starsExtra::make_grid(peatlandDelimination, 1)
  myGrid <- sf::st_crop(myGrid, peatlandDelimination)
  
  for(i in powerRange){
    
    # Get the MAE
    temp2 <- krige.cv(Dybde ~ 1, peatDepths, set = list(idp=i), nmax = nmax)
    temp$MAE[temp$power==i] <- mean(abs(temp2$residual))
    
    #  Get the volume
    vol_temp <- gstat::idw(Dybde ~ 1, peatDepths, 
                    newdata=myGrid, 
                    nmax=nmax, 
                    idp=i)
    
    vol <- c(vol, sum(vol_temp$var1.pred, na.rm=T))
}
  
  
  ifelse(temp$power[which.min(temp$MAE)]<2,
  temp$best <- ifelse(temp$power==2, "best", "not-best"),
  temp$best <- ifelse(temp$MAE==min(temp$MAE), "best", "not-best")
  )

  # Plot MAE
  gg_out <- ggplot(temp, aes(x = power, y = MAE,
                      colour = best,
                      shape = best))+
  geom_point(size=10)+
  theme_bw(base_size = 12)+
  scale_x_continuous(breaks = powerRange)+
  guides(colour="none",
         shape = "none")+
  scale_color_manual(values = c("darkgreen","grey"))+
  scale_shape_manual(values = c(18, 19))+
  ggtitle(title)
  
  # Plot volume
  vol_df <- data.frame("volume" = vol,
                              "power" = powerRange)
  vol_df$relative_volume <- vol_df$volume/mean(vol_df$volume)*100

  
  gg_out_vol <- 
    ggplot(vol_df, aes(x = factor(power), y = relative_volume))+
       geom_point(size=8)+
       xlab("power")+
       ylab("Peat volume as a percentage of\nmean predicted peat volume")+
       theme_bw(base_size = 12)
  
  ggpubr::ggarrange(gg_out,
                    gg_out_vol)
}



```

```{r}
ccalc_optimumPower(peatDepths = depths_tydal_i60,
                   title = "Tydal",
                   peatlandDelimination = SHP_tydal)
```

That's good. The optimum power is the same as when we had all the data points.
