# Carbon stocks

Now I want to convert the peat volumes into estimates of carbon stock. This conversion depends on the values of some peat characteristics, such as bulk density and C concentrations. We have a data set with these values.

```{r}
peatCharacteristics <- read_delim("Data/peatCharacteristics.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
head(peatCharacteristics)
```

```{r}
# Each sample consists of multiple sub samples from different depths. Here we will ignore the depth aspect, and simply take te mean from each sample core.
peatCharacteristics_summedDepths <- peatCharacteristics %>% 
      mutate(perc_SOM = as.numeric(`% SOM`)) %>% 
      mutate(BD = as.numeric(`BD (t/m3)`)) %>% 
      dplyr::select(`SAMPLE ID2`, perc_SOM, BD) %>% 
      group_by(`SAMPLE ID2`) %>% 
      summarise(across(.fns = list(mean =~ mean(., na.rm=TRUE))))

# Getting the other variables that I also want to keep
df_info <- peatCharacteristics %>% 
      dplyr::select(`SAMPLE ID2`, `General Peatland Type`, `Specific Peatland Type`) %>% 
      group_by(`SAMPLE ID2`) %>% 
      unique()
# and join together again (this two-stap proceedure could be simplified)
df = full_join(peatCharacteristics_summedDepths, df_info, by = "SAMPLE ID2") %>% drop_na()
head(df)
```

```{r}
table(df$`General Peatland Type`)
```
We have 12 samples from bogs, and 14 from fens.

```{r}
table(df$`Specific Peatland Type`)
```
We do not perhaps have enough data points to really compare the different spesific peatland types. 

Lets get the total peat volumes for Tydal (and Geilo, but I will not look at Geilo here).
```{r}
volume
```

Calculating the c stocks (tons C) for the Tydal site without specifying the mire type
```{r}
c_stock_tydal_uninformed <- 
              volume$volume[volume$site=="Tydal"] * 
              mean(df$perc_SOM_mean / 100) * 
              mean(df$BD_mean) * 0.5 #* 1000 # removing the original conversion from tons to kg
c_stock_tydal_uninformed
```
I think we also need to estimate the uncertainties around this number. Note that for the peat volume estimation, we did not calculate an uncertainty. But for the peat characteristics we have an uncertainty in therms of variation within and between samples points. I will calculate the uncertainty between points only.
```{r, fig.cap="Estaimted carbon stock in the Tydal peatland"}
c_stock_tydal_uninformed <- NULL
for(i in 1:1000){
 temp <-      volume$volume[volume$site=="Tydal"] * 
              mean(sample(df$perc_SOM_mean, replace = T) / 100) * 
              mean(sample(df$BD_mean, replace = T) * 0.5)
 
 c_stock_tydal_uninformed <- c(c_stock_tydal_uninformed, temp)
  }
hist(c_stock_tydal_uninformed, main="",
     xlab="C stock")
```

This distribution is quite wide. Let's summarize the distribution.
```{r}
c(quantile(c_stock_tydal_uninformed, c(.05, .5, .95)),
  mean = mean(c_stock_tydal_uninformed), sd = sd(c_stock_tydal_uninformed))
```

We can put this stuff into a more generic function.
```{r ccalc_cStocks}
ccalc_cStocks <- function(volume,
                          peatlandType = c("fen", "bog"),
                          peatData){
  temp_stocks <- NULL
  temp_peatData <-  peatData[peatData$`General Peatland Type` %in% peatlandType,]
    
  for(i in 1:1000){
    temp <-   volume * 
              mean(sample(peatData$perc_SOM_mean, replace = T) / 100) * 
              mean(sample(peatData$BD_mean, replace = T) * 0.5)
 
    temp_stocks <- c(temp_stocks, temp)
  }
  return(c(quantile(temp_stocks, c(.05, .5, .95)),
  mean = mean(temp_stocks), sd = sd(temp_stocks)))
}
```


Calculating summary statistics for the carbon stocks in Tydal, using both general peatland types. 
```{r}
(cstocks_tydal_unspecified <- 
  ccalc_cStocks(volume = volume$volume[volume$site=="Tydal"],
              peatData = df))
```

And using only  the *bog* data.
```{r}
(cstocks_tydal_bog <- 
  ccalc_cStocks(volume = volume$volume[volume$site=="Tydal"],
              peatData = df,
              peatlandType = "bog"))
```

And finally, aslo the fen data.
```{r}
(cstocks_tydal_fen <- 
  ccalc_cStocks(volume = volume$volume[volume$site=="Tydal"],
              peatData = df,
              peatlandType = "fen"))
```



```{r}
cstocks_tydal_unspecified_df <- as.data.frame(cstocks_tydal_unspecified)
names(cstocks_tydal_unspecified_df) <- "C stocks"
cstocks_tydal_unspecified_df$summary <- row.names(cstocks_tydal_unspecified_df)
cstocks_tydal_unspecified_df$information <- "Unspesified"


cstocks_tydal_bog_df <- as.data.frame(cstocks_tydal_bog)
names(cstocks_tydal_bog_df) <- "C stocks"
cstocks_tydal_bog_df$summary <- row.names(cstocks_tydal_bog_df)
cstocks_tydal_bog_df$information <- "Bog"

cstocks_tydal_fen_df <- as.data.frame(cstocks_tydal_fen)
names(cstocks_tydal_fen_df) <- "C stocks"
cstocks_tydal_fen_df$summary <- row.names(cstocks_tydal_fen_df)
cstocks_tydal_fen_df$information <- "Fen"

cstocks_tydal_compare <- rbind(
  cstocks_tydal_fen_df,
  cstocks_tydal_bog_df,
  cstocks_tydal_unspecified_df
)

# Pivot
cstocks_tydal_compare <- pivot_wider(cstocks_tydal_compare,
                                      names_from = summary,
                                      values_from = "C stocks"
                                      )
```

```{r, fig.cap="Mean (+- 95 CI) carbon stock for the Tydal test site."}
ggplot(cstocks_tydal_compare,
       aes(x = information))+
  geom_point(aes(y = mean), shape=15, size=10) +
  geom_linerange(aes(ymin=`5%`, 
                     ymax=`95%`),
                 size=2)+ 
  theme_bw(base_size = 16)+
  labs(x = "Peat characteristics",
       y = "Carbon stocks (tons)")
```

Specifying the peatland type makes no difference!

Let's just get the (non-informed) C stock estimates for Geilo as well
```{r}
ccalc_cStocks(volume = 102444.14,
              peatData = df)
```


