# Additional test sites

We used two contrastong sites, Tydal and Geilo, for the main analyses and testing. Now I will bring in an additinal three sites to validate the gernerality of these finding.

## Modalen
Import shape files with peatland delineations.
```{r, message=F, warning=F}
SHP_modalen <- sf::read_sf("Data/Modalen/Modalen_myrareal.shp")
#st_crs(SHP_modalen) #OK
```

Another shape filw with the project area (see below)
```{r}
SHP_modalen_project <- sf::read_sf("Data/Modalen/modalen_shp.shp") 
#st_crs(SHP_modalen_project) # OK
```


Import depth samples
```{r}
depths_modalen <- sf::read_sf("Data/Modalen/modalen_punkter.shp")
#st_crs(depths_modalen) #OK

```


```{r, fig.cap="The Modalen test site sith peatland delineated as grey polygons. Crosses are peat depth samples, contained with in the project area (dotted line)."}
tm_shape(SHP_modalen)+
  tm_polygons()+
  tm_shape(depths_modalen)+
  tm_symbols(col="black",
             size=.5,
             shape=4)+
  tm_shape(SHP_modalen_project)+
  tm_borders(lty="dotted")+
  tm_compass(type="8star", position = c("right", "bottom"), size = 2) +
  tm_scale_bar(position = c("left", "bottom"), width = 0.3)+
  tm_layout(inner.margins = c(0.15, 0.05, 0.05, 0.05))
```
The depth measurements are done within a bigger project area, also covering non-peat areas. The depth measurements come from ground boring, rather than from a simple peat probe, and may reflect the depth including some mineral layers at the bottom. We therefore should not use those points in the IDW I think. This example illustrates the importance of appropriate sampling. We are now limited to just a few data points. The smaller mires have only one or two datapoints.   

Calculating the combined area of the five mires. 
```{r}
SHP_modalen$myArea <- sf::st_area(SHP_modalen)
sum(SHP_modalen$myArea)
```

And the area of the project delineation
```{r}
SHP_modalen_project$myArea <- sf::st_area(SHP_modalen_project)
SHP_modalen_project$myArea
```

Removing the depth measurement from outside the peatland delinitaion
```{r}
depths_modalen_reduced <- sf::st_intersection(depths_modalen, SHP_modalen)
```

```{r, fig.cap="The Modalen test site sith peatland delineated as grey polygons. Crosses are peat depth samples, contained with in the project area (dotted line). Only depth measurements that intersects the mire polygons are included"}
tm_shape(SHP_modalen)+
  tm_polygons()+
  tm_shape(depths_modalen_reduced)+
  tm_symbols(col="black",
             size=.5,
             shape=4)+
  tm_shape(SHP_modalen_project)+
  tm_borders(lty="dotted")+
  tm_compass(type="8star", position = c("right", "bottom"), size = 2) +
  tm_scale_bar(position = c("left", "bottom"), width = 0.3)+
  tm_layout(inner.margins = c(0.15, 0.05, 0.05, 0.05))
```

Create a raster grid
```{r}
#First fix polygon closure
SHP_modalen <- sf::st_make_valid(SHP_modalen)


grid_Modalen_stars_crop <- starsExtra::make_grid(SHP_modalen, 1) %>%
  sf::st_crop(SHP_modalen)
```

```{r, fig.cap="Checking that the raster grid for Modalen looks correct."}
tm_shape(grid_Modalen_stars_crop)+
  tm_raster()
```

Predict peat depth
```{r}
ccalc_optimumPower(nmax=nrow(depths_modalen_reduced),
                   peatDepths = depths_modalen_reduced,
                   title="Modalen",
                   peatlandDelimination = SHP_modalen)
```

Unsurprising, when we have so few data points, the optimum power is very low. I will predict using power 1 through 3 to compare, just because I don't trust power 1 is the best choice. 

```{r}
IDW_Modalen_1 <- gstat::idw(formula = Dybde ~ 1, 
           locations = depths_modalen_reduced, 
           newdata = grid_Modalen_stars_crop, 
           idp=1,
           nmax = 20) # nmax > than the number of points

IDW_Modalen_2 <- gstat::idw(formula = Dybde ~ 1, 
           locations = depths_modalen_reduced, 
           newdata = grid_Modalen_stars_crop, 
           idp=2,
           nmax = nrow(depths_modalen_reduced))

IDW_Modalen_3 <- gstat::idw(formula = Dybde ~ 1, 
           locations = depths_modalen_reduced, 
           newdata = grid_Modalen_stars_crop, 
           idp=3,
           nmax = nrow(depths_modalen_reduced))
```

```{r, fig.cap="IDW for peat depth at Modalen test site."}
tmap_arrange(
tm_shape(IDW_Modalen_1)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 1\nInterpolated peat\ndepth (m)",
            breaks = seq(0,3,.5))+
  tm_compass(type="8star", position = c("right", "bottom"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), width = 0.3)+
  tm_shape(depths_modalen_reduced)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
,
tm_shape(IDW_Modalen_2)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 2\nInterpolated peat\ndepth (m)",
            breaks = seq(0,3,.5))+
  tm_compass(type="8star", position = c("right", "bottom"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), width = 0.3)+
  tm_shape(depths_modalen_reduced)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
,
tm_shape(IDW_Modalen_3)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 3\nInterpolated peat\ndepth (m)",
            breaks = seq(0,3,.5))+
  tm_compass(type="8star", position = c("right", "bottom"), size = 2) +
  tm_scale_bar(position = c("right", "bottom"), width = 0.3)+
  tm_shape(depths_modalen_reduced)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)

)
  
```

I think power 3 looks more accurate because of the way it gives weight to local points. The recommended practice for this case I think would be to analyse each polyong or mire separately because of how the point from neighboring mires affect the depth predictions.

Summary stats for the peat depth at Modalen.
```{r}
summary(depths_modalen_reduced$Dybde)
```

The MAE for the power =3 prediction is 0.82 m.

The rough volume (in m3), that is the mean peat depth multiplied by the area, is
```{r}
mean(depths_modalen_reduced$Dybde)*sum(SHP_modalen$myArea)
```
Number of depth samples per 100 m2:
```{r}
12/(sum(SHP_modalen$myArea)/100)
```

Interpolated volume
```{r}
(volume_modalen <- sum(IDW_Modalen_3$var1.pred, na.rm=T))
```

Estimated carbon stocks for the Modalen site:
```{r}
ccalc_cStocks(volume_modalen, peatData = df)
```



## Opelandsmarka
Import shape file
```{r import shape files, message=F, warning=F}
SHP_opelandsmarka <- sf::read_sf("Data/Opelandsmarka/Opelandsmarka Voss.shp")
#st_crs(SHP_opedalsmarka) #32632 UTM 32
```

The depth measurements are in three different files.
```{r, message=F, warning=F}
depths_opelandsmarka1 <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr1.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
depths_opelandsmarka2 <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr2.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
depths_opelandsmarka3 <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr3.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
# The last one is missing a column

depths_opelandsmarka <- bind_rows(depths_opelandsmarka1,
                                  depths_opelandsmarka2,
                                  depths_opelandsmarka3)


depths_opelandsmarka <- sf::st_as_sf(depths_opelandsmarka,
                                     coords = c("UTM_N_ZONE32N", "UTM_E_ZONE32N"),
                                     crs = 32632)
```




```{r, fig.cap="Test site: Opelandsmarka. Crosses are peat depth measurements."}
tm_shape(SHP_opelandsmarka)+
  tm_polygons()+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```

I think we can ignore that top mire as it was not sampled nearly as intensive as the other two.
```{r}
SHP_opelandsmarka_twoMires <- SHP_opelandsmarka[-1,]
```

It could be a good idea sometimes to include depth measurements outside the mire polygons (with a depth of 0). Here it is easier for me to just exclude all points outide any polygon.
```{r}
depths_opelandsmarka <- st_intersection(depths_opelandsmarka, SHP_opelandsmarka_twoMires)
```
```{r, fig.cap="Opelandsmarka with two out of three mires. Crosses are peat depth measurements."}
tm_shape(SHP_opelandsmarka_twoMires)+
  tm_polygons()+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```

There is one NA in the depth data which we have to remove
```{r}
depths_opelandsmarka <- depths_opelandsmarka[!is.na(depths_opelandsmarka$Dybde),]
```

Try again:

```{r, fig.cap="Determening the optimal power for the Opelandsmarka test site."}
ccalc_optimumPower(peatDepths = depths_opelandsmarka, 
                   title = "Opelandsmarka",
                   peatlandDelimination = SHP_opelandsmarka_twoMires)
```

Again the lowest MAE is obtained from a power of 1, but as we have seen, the power should probably be set higher. The predicted peat volume for example is quite unstable at low power settings.

Create a raster grid for the predictions
```{r}
grid_Opelandsmarka_stars_crop <- starsExtra::make_grid(SHP_opelandsmarka_twoMires, 1) %>%
  sf::st_crop(SHP_opelandsmarka_twoMires)
```

The depths are recorded as cm not meters. Converting to m. 
```{r}
depths_opelandsmarka$Dybde <- depths_opelandsmarka$Dybde/100
```


```{r}
IDW_opelandsmarka_1 <- gstat::idw(Dybde ~ 1, depths_opelandsmarka, 
                    newdata=grid_Opelandsmarka_stars_crop, 
                    nmax=nmax, 
                    idp=1)

IDW_opelandsmarka_2 <- gstat::idw(Dybde ~ 1, depths_opelandsmarka, 
                    newdata=grid_Opelandsmarka_stars_crop, 
                    nmax=nmax, 
                    idp=2)

IDW_opelandsmarka_3 <- gstat::idw(Dybde ~ 1, depths_opelandsmarka, 
                    newdata=grid_Opelandsmarka_stars_crop, 
                    nmax=nmax, 
                    idp=3)
```

```{r}
myBreaks <- seq(0,3,.5)
myPos <- c("left", "bottom")

tmap_arrange(
tm_shape(IDW_opelandsmarka_1)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 1\nInterpolated peat\ndepth (m)",
            breaks = myBreaks)+
  tm_compass(type="8star", position = myPos, size = 2) +
  tm_scale_bar(position = myPos, width = 0.3)+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
,
tm_shape(IDW_opelandsmarka_2)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 2\nInterpolated peat\ndepth (m)",
            breaks = myBreaks)+
  tm_compass(type="8star", position = myPos, size = 2) +
  tm_scale_bar(position = myPos, width = 0.3)+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
,
tm_shape(IDW_opelandsmarka_3)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 3\nInterpolated peat\ndepth (m)",
            breaks = myBreaks)+
  tm_compass(type="8star", position = myPos, size = 2) +
  tm_scale_bar(position = myPos, width = 0.3)+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T),
ncol = 2

)
```

With power set to 1 the model is too flat, and the deep areas are not identified correctly. Power 2 or 3 makes less of a difference. Becaus ethe MAE and volume estimates are so similar, I will chose the modet local model (power = 3)

Getting the total peatland area
```{r}
SHP_opelandsmarka_twoMires$myArea <- sf::st_area(SHP_opelandsmarka_twoMires)
sum(SHP_opelandsmarka_twoMires$myArea)
```

Getting the mean peat depth (after excluding points outside the polygons)
```{r}
mean(depths_opelandsmarka$Dybde)
```

Rough volume
```{r}
mean(depths_opelandsmarka$Dybde) * sum(SHP_opelandsmarka_twoMires$myArea)
```

Number of peat depth measurements
```{r}
nrow(depths_opelandsmarka)
```

Depth measurement density (points per 100 m-2)
```{r}
nrow(depths_opelandsmarka)/(sum(SHP_opelandsmarka_twoMires$myArea)/100)
```

Interpolated peat volume (best model)
```{r}
(vol_temp <- sum(IDW_opelandsmarka_3$var1.pred, na.rm=T))
```

Estimated C stocks
```{r}
ccalc_cStocks(volume = vol_temp,
              peatData = df)
```

## Kinn 1

```{r}
# import peatland delineation
SHP_kinn1 <- sf::st_read("Data/Kinn/rikmyr_avgrensing.shp")

# import peat deth data
depths_kinn1 <- read_delim("Data/Kinn/data_rikmyr_torvdybder.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
```

The peat depth data doesn't contain coordinated. Importing those now.
```{r}
depths_kinn1_spatial <- sf::st_read("Data/Kinn/rikmyr_punkter.shp")

```

The column `name` is shared between the two datasets so I can use this to merge them. First I need to remove leading zeros.
```{r}
depths_kinn1_spatial$name <- as.numeric(depths_kinn1_spatial$name)
```


```{r}
depths_kinn1_spatial$Dybde <- depths_kinn1$Dybde[match(depths_kinn1_spatial$name, depths_kinn1$name)]
rm(depths_kinn1)
```

There is still one NA to remove
```{r}
depths_kinn1_spatial <- depths_kinn1_spatial[!is.na(depths_kinn1_spatial$Dybde),]
```



Check CRS
```{r}
#st_crs(depths_kinn1_spatial) # lat long
#st_crs(SHP_kinn1) # also lat long

depths_kinn1_spatial <- sf::st_transform(depths_kinn1_spatial, 25833)
SHP_kinn1 <- sf::st_transform(SHP_kinn1, 25833)
```

```{r, fig.cap="Test site Kinn1."}
tm_shape(SHP_kinn1)+
  tm_polygons()+
  tm_shape(depths_kinn1_spatial)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```

This is a nice example data set. I notice however, there are few points near the edges.

Getting the rough volume
```{r}
SHP_kinn1$myArea <- sf::st_area(SHP_kinn1)
SHP_kinn1$myArea
```

Mean depth
```{r}
mean(depths_kinn1_spatial$Dybde, na.rm=T)
```

The depths are in cm. Converting to m.
```{r}
depths_kinn1_spatial$Dybde <- depths_kinn1_spatial$Dybde/100
```

Number of depth measurements
```{r}
nrow(depths_kinn1_spatial)
```

Point density (100 m-2)
```{r}
nrow(depths_kinn1_spatial) / (SHP_kinn1$myArea )*100

```

Finding optimum power
```{r, fig.cap="Finding the best power setting for the Kinn1 test site."}
ccalc_optimumPower(peatDepths = depths_kinn1_spatial,
                   peatlandDelimination = SHP_kinn1,
                   title = "Kinn1")
```

The optimal power is 5.

Create a raster grid for the predictions
```{r}
grid_kinn1_stars_crop <- starsExtra::make_grid(SHP_kinn1, 1) %>%
  sf::st_crop(SHP_kinn1)
```

```{r}
IDW_kinn1_5 <- gstat::idw(Dybde ~ 1, depths_kinn1_spatial, 
                    newdata=grid_kinn1_stars_crop, 
                    nmax=nmax, 
                    idp=5)
```

Max peat depth
```{r}
max(depths_kinn1_spatial$Dybde)
```


```{r, fig.cap="Best model for test site Kinn1"}
myBreaks <- seq(0,5,1)
myPos <- c("right", "bottom")

tm_shape(IDW_kinn1_5)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 5\nInterpolated peat\ndepth (m)",
            breaks = myBreaks)+
  tm_compass(type="8star", position = myPos, size = 2) +
  tm_scale_bar(position = myPos, width = 0.3)+
  tm_shape(depths_kinn1_spatial)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
```

Predicted peat volume
```{r}
(vol_temp <- sum(IDW_kinn1_5$var1.pred, na.rm=T))
```

Carbon stocks
```{r}
ccalc_cStocks(volume = vol_temp,
              peatData = df)
```

## Kinn2

```{r}
SHP_kinn2 <- sf::st_read("Data/Kinn/maroy_avgrensing.shp")
```


```{r}
depths_kinn2 <- read_delim("Data/Kinn/data_maroy_torvdybder.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
```

As before, the depths have no coordinates.
```{r}
depths_kinn2_spatial <- sf::st_read("Data/Kinn/maroy_punkter.shp")
```

```{r}
depths_kinn2_spatial$name <- as.numeric(depths_kinn2_spatial$name)
depths_kinn2_spatial$Dybde <- depths_kinn2$Dybde[match(depths_kinn2_spatial$name, depths_kinn2$name)]
rm(depths_kinn2)
```


Transform to UTF33
```{r}
depths_kinn2_spatial <- sf::st_transform(depths_kinn2_spatial, 25833)
SHP_kinn2 <- sf::st_transform(SHP_kinn2, 25833)
```



```{r, fig.cap="Test site Kinn2."}
tm_shape(SHP_kinn2)+
  tm_polygons()+
  tm_shape(depths_kinn2_spatial)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```

The mean  depth
```{r}
summary(depths_kinn2_spatial$Dybde, na.rm=T)
```

Converting from cm to m
```{r}
depths_kinn2_spatial$Dybde <- depths_kinn2_spatial$Dybde/100
```


Area
```{r}
SHP_kinn2$myArea <- st_area(SHP_kinn2)
nrow(depths_kinn2_spatial)
```

Number of peat depth measurements
```{r}
nrow(depths_kinn2_spatial)
```

Density of points (100 m-2)
```{r}
nrow(depths_kinn2_spatial)/SHP_kinn2$myArea*100
```

Find optimal power
```{r, fig.cap="Finding the best power setting for the Kinn2 test site."}
ccalc_optimumPower(peatDepths = depths_kinn2_spatial,
                   peatlandDelimination = SHP_kinn2,
                   title = "Kinn2")
```

Power = 4 is best.

Create a raster grid for the predictions
```{r}
grid_kinn2_stars_crop <- starsExtra::make_grid(SHP_kinn2, 1) %>%
  sf::st_crop(SHP_kinn2)
```

Interpreted peat volume
```{r}
IDW_kinn2_4 <- gstat::idw(Dybde ~ 1, depths_kinn2_spatial, 
                    newdata=grid_kinn2_stars_crop, 
                    nmax=nmax, 
                    idp=4)

```

```{r, fig.cap="Best model for test site Kinn2"}
myBreaks <- seq(0,3,.5)
myPos <- c("right", "bottom")

tm_shape(IDW_kinn2_4)+
  tm_raster(col="var1.pred",
            palette = "-viridis",
            title = "Power = 4\nInterpolated peat\ndepth (m)",
            breaks = myBreaks)+
  tm_compass(type="8star", position = myPos, size = 2) +
  tm_scale_bar(position = myPos, width = 0.3)+
  tm_shape(depths_kinn2_spatial)+
  tm_symbols(shape=4,
             col="black",
             size=.5)+
  tm_layout(legend.outside = T)
```

And the volume is
```{r}
(vol_temp <- sum(IDW_kinn2_4$var1.pred, na.rm=T))
```

And the C stock
```{r}
ccalc_cStocks(volume = vol_temp,
              peatData = df)
```

