# Additional test sites

We used two contrastong sites, Tydal and Geilo, for the main analyses and testing. Now I will bring in an additinal three sites to validate the gernerality of these finding.

## Modalen
Import shape files with peatland delineations.
```{r import shape files, message=F, warning=F}
SHP_modalen <- sf::read_sf("Data/Modalen/Modalen_myrareal.shp")
#st_crs(SHP_modalen) #OK
```

Import depth samples
```{r}
depths_modalen <- sf::read_sf("Data/Modalen/modalen_punkter.shp")
#st_crs(depths_modalen) #OK

```


```{r, fig.cap="The Modalen test site sith peatland delineated as grey polygons. Crosses are peat depth samples."}
tm_shape(SHP_modalen)+
  tm_polygons()+
  tm_shape(depths_modalen)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```

Im not sure how to approach this. Why is there peat depth data from outside the peatland?

## Opelandsmarka

```{r import shape files, message=F, warning=F}
SHP_opelandsmarka <- sf::read_sf("Data/Opelandsmarka/Opelandsmarka Voss.shp")
#st_crs(SHP_opedalsmarka) #32632 UTM 32

```


```{r, eval=F}
depths_opelandsmarka <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr1.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
depths_opelandsmarka2 <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr2.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)
depths_opelandsmarka3 <- read_delim("Data/Opelandsmarka/Torvdybder_Opelandsmarka_myr3.csv", 
    delim = ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"), 
    trim_ws = TRUE)


depths_opelandsmarka <- sf::st_as_sf(depths_opelandsmarka3,
                                     coords = c("UTM_N_ZONE32N", "UTM_E_ZONE32N"),
                                     crs = 32632)
```
Not clear which one to use here. I will focus on just the middle peatland.

```{r}
tm_shape(SHP_opelandsmarka)+
  tm_polygons()+
  tm_shape(depths_opelandsmarka)+
  tm_symbols(col="black",
             size=.5,
             shape=4)
```



