---
title: "Interpolation_Tydal_20220122"
author: "MarteFandrem"
date: "20 1 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Packages needed
```{r, echo=FALSE} 
library(readxl)
library(writexl)
library(rgdal)
library(raster)
library(ggplot2)
library(gstat)
library(sf)
library(broom)
library(ggthemes)
library(viridis)
library(sp)
library(spatialEco)
library(spm)
library(tmap)
library(Metrics)
library(rlist)
library(dplyr)
```

#Import and clean up data

```{r}

shp<-readOGR(dsn="C:/Users/martef/OneDrive - NTNU/Documents/Jobb-pc/PhD prosjekter/WP3 NINA GRAN/Data/Rydda datasett for kalkulator/Tydal", layer="stasjon_Setermyra")

df <- read.csv("C:/Users/martef/OneDrive - NTNU/Documents/Jobb-pc/PhD prosjekter/WP3 NINA GRAN/Data/Rydda datasett for kalkulator/Tydal/Torvdybder_Tydal.csv", sep=";")

# Make spatial dataframe of the peat depths points

dfs <- st_as_sf(x = df, 
                        coords = c("x", "y"),
                        crs = "+init=epsg:25833")
#GDAL Message 1: +init=epsg:XXXX syntax is deprecated. It might return a CRS with a non-EPSG compliant axis order.
#Don't know how to write it otherwise. Seems to work properly though.

sf_shp <- st_as_sf(shp)

sf::st_crs(shp)
sf::st_crs(dfs)

proj4string(shp)<-crs(dfs)

dfsp <- as(dfs, Class="Spatial")

#Crop dfsp to the extent of the station area
dfsp_station <- dfsp[shp,]
```

#Visualisere dataene

```{r}
tmap_mode("view")
  tm_shape(sf_shp)+
  tm_polygons() +
  tm_shape(dfs)+
  tm_dots(col="black", size=0.01, alpha=0.5, )

```

#Create grid and adjust to extent of peatland

```{r}
grid <- raster(extent(shp)) #create a raster grid from the extent of the peatland
res(grid) <- 1              #set resolution of the grid to 1x1m
proj4string(grid)<-crs(dfs) #set similar projection to the grid as to the datapoints

grid_sp <-as(grid, "SpatialPixels") #convert the grid from raster to spatialpixels

grid_sp@grid@cellsize       #check that cell size is 1x1

grid_crop <- grid_sp[shp,]  #crop the grid to only include the peatland
plot(grid_crop)


```

Setup functions for running LOOCV and interpolations

```{r}
cv.IDW  <- function(dfsp_station, stat.formula = NULL,
                    seqNeighbors = NULL, seqBeta = NULL,
                    evalGridSize = NULL, 
                    evalRaster = NULL, 
                    verbose = TRUE){}


```

```{r}
RMSE <- function(residuals){
  sqrt(sum((residuals)^2)/length(residuals))
}
```


Provide default values for function arguments
```{r}

  if (is.null(seqNeighbors)){
    seqNeighbors <- round(seq(2, length(dfsp_station), by = 2))
  }
  if (is.null(seqBeta)){
    seqBeta <- c(0.1, seq(0.5, 6, 0.5))
  }
  if (is.null(evalGridSize)){
    x.interval <- extent(shp)@xmax - extent(shp)@xmin
    y.interval <- extent(shp)@ymax - extent(shp)@ymin
    evalGridSize <- round(min(x.interval, y.interval) *0.005)
  }
  if (is.null(stat.formula)){
    stat.formula=Dybde~1
    
  }
  if (is.null(evalRaster)){
    extent.evalGrid <- extent(shp)
  }else{
    extent.evalGrid <- extent(evalRaster)
  }
  
  
```

Build a grid for parameter combinations
```{r}
  cv.Grid <- expand.grid(Beta = seqBeta,
                         Neighbors = seqNeighbors)
  cv.Grid$RMSE <- NA
```

Perform IDW and LOOCV
```{r}
 for (i in 1:nrow(cv.Grid)){
    ### BUILD IDW MODEL ###
    idw <- gstat(formula = Dybde~1,
                data = dfsp_station, 
                nmax = cv.Grid[i, 'Neighbors'], 
                set = list(idp = cv.Grid[i, 'Beta']))
    ### PERFORM LOOCV ###
    crossval <- gstat.cv(idw, 
                         nmax = cv.Grid[i, 'Neighbors'],
                         beta = v.Grid[i, 'Beta'],
                         debug.level = 0)
    cv.Grid[i, 'RMSE'] <- RMSE(crossval$residual)
    if (verbose){
      print(paste('Function call', i, 'out of',  nrow(cv.Grid)))
      print(paste('Evaluating beta =', 
                  cv.Grid[i, 'Beta'], 
                  'and neighbors =',  
                  cv.Grid[i, 'Neighbors']))
      print(paste('RMSE=', RMSE(crossval$residual)))
    }
  }

```
```{r}
ggplot(cv.Grid, aes(Neighbors, Beta)) + 
  geom_tile(aes(fill = RMSE), colour = "black") + 
  scale_fill_gradient(low = "steelblue", high = "orange") + 
  theme_bw() +
  ggtitle('Parameter values')
```



### LOOP THROUGH ALL PARAMETER COMBINATIONS ###
  for (i in 1:nrow(cv.Grid)){
    ### BUILD IDW MODEL ###
    idw <- gstat(formula = stat.formula,
                data = spatialDF, 
                nmax = cv.Grid[i, 'Neighbors'], 
                set = list(idp = cv.Grid[i, 'Beta']))
